BioTime

### 2019-04-16 biotime.Rmd
Get duplicate id's via function. Get complete trait data with id's.

Some names match to multiple acceptedUsage ID's, so we have to manually choose which one we want. In this case it's just the Blue-winged warbler
```{r}
unres <- biotime_ids %>%
  distinct(acceptedNameUsageID, input, sort) %>%
  group_by(sort) %>% 
  filter(n()>1) %>% 
  ungroup()

unres
```

Add the unresolved ID to all the other ids and join with biotime data
```{r}
biotime_ids <- unres %>%
left_join(biotime_ids) %>% #join observations back with all the columns
  filter(taxonomicStatus == "accepted") %>% #want the accepted ID's
  bind_rows(biotime_ids %>% filter(!sort %in% unres$sort)) #join to resolved ID's, excluding the observations we resolved manually

@@

##Elton Traits

```{r}
bird_ids <- by_name(unique(bird_trait$scientific), "itis") %>%
  bind_rows(by_common(unique(bird_trait$scientific), "itis")) %>%
  #filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

check if all bird biotime data w/id's has an elton match. There are 54 species without a match
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id) %>%
  distinct() %>%
  filter(!id %in% bird_ids$acceptedNameUsageID) %>%
  dim()
```

and 605 entries in the species column without an ID
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id, genus_species) %>%
  distinct() %>%
  filter(is.na(id)) %>%
  dim()
```

>>>>

Some names match to multiple acceptedUsage ID's, so we have to manually choose which one we want. In this case it's just the Blue-winged warbler
```{r}

get_dupe_ids <- function(data){
  biotime_ids %>%
    distinct(acceptedNameUsageID, input, sort) %>%
    group_by(sort) %>% 
    filter(n()>1) %>% 
    ungroup()
}

unres <- get_dupe_ids(biotime_ids)
unres 
```

Add the unresolved ID to all the other ids and join with biotime data
```{r}
biotime_ids <- unres %>%
left_join(biotime_ids) %>% #join ID's back with all the columns
  filter(taxonomicStatus == "accepted") %>% #want the accepted ID from the two Warbler options
  bind_rows(biotime_ids %>% filter(!sort %in% unres$sort)) #join to resolved ID's, excluding the observations we resolved manually

@@

##Elton Traits
Get id's for the Elton Bird trait data
```{r}
bird_ids <- by_name(unique(bird_trait$scientific), "itis") %>%
  bind_rows(by_common(unique(bird_trait$scientific), "itis")) %>%
  #filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

Are there any unresolved ids? Yes - just one, which is a synonym to two different accepted id's
```{r}
unres_trait <- get_dupe_ids(bird_ids) %>%
  left_join(bird_ids)

unres_trait

#resolve to the desired ID (in this case the one with complete hierarchy data)
bird_ids <- unres_trait %>% 
  filter(!is.na(kingdom)) %>%
  bind_rows(bird_ids %>% filter(!sort %in% unres_trait$sort))

bird_trait_data <- bird_ids  %>%
  select(acceptedNameUsageID, input) %>%
  distinct() %>%
  right_join(bird_trait, by = c("input" = "scientific")) %>%
  rename(scientific = input)
```

check if all bird biotime data w/id's has an elton match. There are 54 species without a match
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id) %>%
  distinct() %>%
  filter(!id %in% bird_trait_ids$acceptedNameUsageID) %>%
  dim()
```

and 605 species without an ID
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id, genus_species) %>%
  distinct() %>%
  filter(is.na(id)) %>%
  dim()
```

Are there any studies for which we have trait data for all species?
```{r}
bird_data <- biotime_data %>%
  filter(taxa == "Birds") %>%
  left_join(bird_ids)
```
